---
title: "ETC5513 Assignment 1"
author: "Caryn Kesuma"
format:
    html: 
        number-sections: true
    pdf: 
        number-sections: true
execute:
    eval: true
    message: false
    warning: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE) 
```

```{r}
#| echo: false

library(tidyverse)
library(lubridate)
library(ggplot2)
library(kableExtra)
```

## Introduction
Businesses heavily rely on personalised marketing to drive customer purchases. Having a deeper understanding on whether *investing more on marketing eventually leads to increased customer spending* is vital for optimising resource allocation. This report analyses customer-level campaign data to evaluate the effectiveness of marketing spend. The following findings will help inform whether increasing marketing budget is worth the investment.

## Research Question
This report investigates whether *higher marketing spend per customer correlates to higher purchase amounts*. Specifically, it examines the relationship between marketing expenditure and customer purchasing behaviour. The objective is to determine if increased spending results in profitable returns for the business. 

## Dataset Introduction
The dataset contains customer-level information from a personalised marketing campaign. It includes variables such as customer ID, sign up date, marketing spend, number of visits, purchase amount, customer segment, and notes. This data records how much was spent targeting each customer and their resulting purchasing behaviour.

Data cleaning involved includes the following:

- Converting date formats into a consistent and the same structure where `parse_date_time` function from `lubridate` package was used because it automatically handles multiple date-time formats at once;

- Handling missing values in purchase amounts; 

- Segment (Consumer, Enterprise, SMB) were converted into factor variables to be treated as categories in analysis and visualisation; and 

- 'notes' column was presumed as irrelevant for analysis, as it does not provide relevant information on marketing spend or customer purchasing behaviour, so it was removed from the dataset.

```{r}
#| echo: false

data <- read.csv("data/37094378_Caryn Kesuma.csv")

data_cleaned <- data %>% 
    mutate(
        signup_date = na_if(signup_date, "NULL"),
        signup_date = na_if(signup_date, "N/A"),
        signup_date = parse_date_time(signup_date, orders = c("dmy", "ymd", "mdy")),
        segment = as.factor(segment)
    ) %>%
    drop_na(purchase_amount) %>%
    select(-notes)
```

### Dataset Description
The raw dataset contains `r nrow(data)` observations and `r ncol(data)` variables.

The cleaned dataset contains `r nrow(data_cleaned)` observations and `r ncol(data_cleaned)` variables.

#### Summary Plot
```{r}
#| label: fig-histogram
#| fig-cap: "Distribution of marketing spend across customers between $0 to $300 across all customers."

ggplot(data_cleaned, aes(x = marketing_spend)) +
    geom_histogram(bins = 20) + 
    labs(title =  "Distribution of Marketing Spend", 
         x = "Marketing Spend", y = "Count")
```
The distribution of marketing spend is shown in @fig-histogram where spending appears relatively evenly spread across a wide range, indicating that customers were targeted with varying levels of marketing investment. This variety provides a fair basis for analysing how differences in marketing spend relate to purchasing behaviour.

#### Summary Table
```{r}
#| label: tbl-summarystats
#| tbl-cap: Summary statistics of the cleaned dataset

data_cleaned %>% summarise(
    avg_spend = mean(marketing_spend, na.rm = TRUE),
    avg_purchase = mean(purchase_amount, na.rm = TRUE),
    avg_visits = mean(visits, na.rm = TRUE)
)
```
Summary statistics in @tbl-summarystats show that the average marketing spend per customer was \$`r round(mean(data_cleaned$marketing_spend), 2)` while the average purchase amount was \$`r round(mean(data_cleaned$purchase_amount), 2)`. Customers made an average of `r round(mean(data_cleaned$visits), 2)` visits. These values suggest moderate engagement and spending levels across the dataset, providing useful context for examining the relationship between marketing investment and customer purchases.

## Results
### Scatter Plot
```{r}
#| label: fig-scatterplot
#| fig-cap: "The relationship between marketing spend and purchase amount."

ggplot(data_cleaned, aes(x = marketing_spend, y = purchase_amount)) +
    geom_point(alpha = 0.6) +
    geom_smooth(method = "lm", se = FALSE) +
    labs(title = "Marketing Spend vs Purchase Amount",
        x = "Marketing Spend", y = "Purchase Amount")
```
@fig-scatterplot shows the relationship between marketing spend and purchasing amount. There is a clear positive trend, where higher marketing spend is associated with higher purchase amounts. This is supported by a strong positive correlation of approximately `r round(cor(data_cleaned$marketing_spend, data_cleaned$purchase_amount), 2)` indicating a strong positive linear relationship between the two variables. On average, a one-unit increase in marketing spend is associated with an increase of approximately $`r round(coef(lm(purchase_amount ~ marketing_spend, data = data_cleaned))[2], 2)` in purchase amount.

The fitted regression line further strengthens this trend, showing the increases in marketing investment are generally associated with a rise in customer spending. While there is still some variability in purchase amounts at each level of marketing spend, the overall pattern remains consistently ascending. This suggests that higher marketing spend per customer is effective in driving increased purchases.

### Boxplot
```{r}
#| label: fig-boxplot
#| fig-cap: "The purchase distribution across the three segments (consumer, enterprise, small business)"
ggplot(data_cleaned, aes(x = segment, y = purchase_amount)) +
    geom_boxplot() +
    labs(title = "Purchase Amount by Customer Segment",
        x = "Segment", y = "Purchase Amount")
```
@fig-boxplot shows the distribution of purchase amounts across different customer segments. While there are some differences in median purchase values between segments, all groups exhibit a wide range of purchase behaviour. This indicates that although customer segment may influence purchasing patterns to some extent, the variation within each segment is notable.

Most importantly, the positive relationship between marketing spend and purchasing amount shown in @fig-scatterplot appears to hold across all the three segments. This suggests that the effect of marketing spend on purchases is consistent and not limited to a specific customer segment.

### Conclusion
Overall, this analysis provides strong evidence that higher marketing spend per customer leads to higher purchasing amounts. The consistent positive relationship observed in the data suggests the effectiveness of increased marketing investment in driving customer spending, and eventually resulting to a successful marketing campaign. 